---
title: "Forest Plot for Stroke Outcome"
author: "Jaskaran Dhiman"
date: "2024-07-19"
output: html_document
---

# Introduction

In this lab, you will learn how to create a forest plot in R. The plot will show the beta estimates for the relationship between diabetes and stroke outcome, adjusted for age, smoking, alcohol, and cholesterol levels.

# Setting Up the Environment

First, we need to load the necessary libraries and the dataset.

```{r}
# install.packages("forestplot")
```

```{r load-libraries}
# Load necessary libraries
library(tidyverse)
library(forestplot)
library(broom)
```

# Overview of the Dataset

We will use a synthetic health dataset that contains information about age, smoking, alcohol consumption, cholesterol levels, diabetes status, and stroke outcome.

```{r load-data}
# Load the dataset
data <- read.csv("health_data.csv")
head(data)
summary(data)
```

The health dataset is a synthetic dataset generated to study the relationship between diabetes and stroke outcomes, adjusted for age, smoking, alcohol consumption, and cholesterol levels. 

Column Descriptions:
Age: The age of the individual (in years).
Smoking: Smoking status (0 = non-smoker, 1 = smoker).
Alcohol: Alcohol consumption status (0 = non-drinker, 1 = drinker).
Cholesterol: Cholesterol level (in mg/dL).
Diabetes: Diabetes status (0 = non-diabetic, 1 = diabetic).
Stroke: Stroke outcome, a continuous variable representing the likelihood of stroke occurrence.


# Performing Linear Regression

Let's perform a linear regression to assess the relationship between diabetes and stroke outcome, adjusted for age, smoking, alcohol, and cholesterol levels.

```{r regression}
# Perform linear regression
model <- lm(Stroke ~ Diabetes + Age + Smoking + Alcohol + Cholesterol, data = data)
summary(model)
```

# Creating the Forest Plot

We will now create a forest plot to visualize the beta estimates from the regression model.

```{r forest-plot}

# Extract coefficients and confidence intervals
coefs <- tidy(model)
coefs <- coefs %>%
  filter(term != "(Intercept)") %>%
  mutate(
    lower = estimate - 1.96 * std.error,
    upper = estimate + 1.96 * std.error
  )


```

```{r}
# Create the forest plot
forestplot(
  labeltext = cbind(coefs$term, round(coefs$estimate, 2), round(coefs$lower, 2), round(coefs$upper, 2)),
  mean = coefs$estimate,
  lower = coefs$lower,
  upper = coefs$upper,
  title = "Forest Plot of Beta Estimates for Stroke Outcome",
  xlab = "Beta Estimate"
)
```

# Conclusion

In this exercise, we covered: - How to perform linear regression in R. - How to create a forest plot to visualize the beta estimates from a regression model.

For further reading, consider exploring more advanced topics like logistic regression and mixed-effects models.

# Additional Resources

-   [R Documentation](https://www.rdocumentation.org/)
-   [forestplot Documentation](https://cran.r-project.org/web/packages/forestplot/forestplot.pdf)
